## The Concept of Democratic Support

There are two major conceptualizations of public support for democracy (PSD)^[Some studies also refer to PSD measured on the national level as "mood".]. First, the "implicit" approach that requires the support for broader sociopolitical values like post-materialism and egalitarianism. Here, people support democracy implicitly if they support the values that are framed as particularly democratic. Second, the "explicit" approach that requires both an appraisal of democracy and a rejection of aut ocratic alternatives. Different studies use different concepts although the explicit approach more direct.

## Drivers of Democratic Support

The main theory about why citizen and societies begin to support democracy are called *generational socialization* and *instrumental regime performance*. The first theory assumes that individuals are taught to support the regime under which they are socialized during late adolescence [@Niemi1974; @Mannheim1970]. One implication is that after a shift to democracy, the support for it will grow over time (@Denemark2016). Indeed, several single-country studies have found evidence for this claim, for example for 1970s Germany (@Baker1981), 1980s Spain (@Montero1997), and 1990s Russia (@Mishler2007). One the other hand, other studies come to different results: First, @Mishler2002 do not find such an effect analyzing other Central and Eastern European countries. Second, more recent studies even find a decline in PSD over the last years [@Foa2016; @Foa2017].

Regarding the second theory, instrumental regime performance, PSD rises if the system performs well in terms of instrumental benefits such as economic growth and it declines if regimes perform poorly [@Dalton1994; @Magalhaes2014]. Consequently, the theory suggests that during PSD declines during economic crises. However, there are case-studies that find [@Dalton1994; @Magalhaes2014; @Mishler1996] and that do not find  find this relationship in the data (@Graham2004).

@Claassen2020mood offers an alternative theory in transferring the thermostatic model of public opinion and policy (@Wlezien1995) to democracy and democratic support. In particular, he suggests that there is a negative feedback loop between PSD and democracy so that PSD decreases if democracy supply increases and vice versa. In short, the reasoning is that the output of democratic rights and institutions overshoots the initial desire for these rights which causes another overcompensation in favor of lower levels of democracy. Moreover, @Claassen2020mood differentiates two causal channels: One for electoral and a second for minoritarian democracies. These two sub-types differ in the degree to which the majority holds juridical power. Therefore, he tests the following hypotheses: First, increases in democracy have a negative effect on PSD (H1), Second, he specifically looks at electoral (H1-elec) and minoritarian democracies (H1-min). In his analysis he finds evidence for H1 and H1-min but not for H1-elec.

## Democratic Support and Survival of Democracy

@Claassen2020support distinguishes between two types of PSD: diffuse and specific. Specific PSD is instrumental and focuses on regime outputs (similar to the instrumental regime performance theory), whereas diffuse support is normative and focuses on the principles of the regime. Therefore, principled PSD is more durable than specific support and helps cushioning regimes in times of political or economic crises. It is thus principled PSD that helps to ensure the survival of democracy. Although the theory is widely accepted (e.g., @Norris2017, @Booth2009, @Mattes2007), so far the findings have also been mixed with supporting contributions [@Inglehart2003; @Welzel2005] and contributions against the theory [@Hadenius2005; @Welzel2007; @Fails2010; @Qi2011]. Astonishingly, these studies all essentially analyze the same data from the World Value Survey starting at wave 3 where respective PSD items are included and still come to different varying results.

@Claassen2020support does not only look at the relationship between PSD and democractic survival but also between PSD and democratic emergence. He includes the argument made by @Qi2011 that PSD may also function as democratic demand, thus increasing the probability of transitioning from autocracy to democracy. Claassen's hypotheses are: First, PSD is positively associated with subsequent change in democracy regardless of the initial level of democracy. Second, he specifically looks at PSD in already-existing democracies (H2-dem) and at PSD in autocracies (H2-aut). He finds evidence in support of H2-dem, mixed evidence for H2 and no evidence for H2-aut.

## Measuring Democratic Support and the Data Challenge

We measure PSD with existing survey data on the national level. In particular, we focus on principled support. The questions ask for the respondent's opinion on the appropriateness or desirability of democracy and undemocratic alternatives and comparisons between both.^[The 8 included survey items and their different variants can be found on [the author's webpage](http://chrisclaassen.com/docs/Democratic_mood_supp_materials.pdf).]

Such items are available in 14 survey projects^[World and European Values Surveys, the Afrobarometer, Arab Barometer, Latinobarometer, Asiabarometer, Asian Barometer, South Asia Barometer, New Europe Barometer, Latin American Public Opinion Project, Eurobarometer, European Social Survey, pew Global Attitudes Project, and the Comparative Study of Electoral Systems], for 150 countries from 1988 to 2017. Combined, the dataset includes 3,765 nationally aggregated binary responses from 1,390 nationally representative survey samples.

The data, however, poses the following challenge: It is highly fractured across time and space, with many -- oftentimes large -- gaps along the time dimension for all countries. Although we look at a 30 years time span, the average number of covered years is slightly below 8. Take, for instance, China as the country with most participants over all years, namely 20895. From 1988 to 2017, however, the data available covers only 8 years: 2001-2003, 2006-2009, and 2011/2012. Many countries are not covered at all or only surveyed once. The South American countries tend to have the best coverage. For example, Argentina tops the list with 23 years included. Germany and the US are available for 12 and 13 years, respectively.

Figure \@ref(fig:sparse-data) gives an overview of the sparseness of the data. The first panel shows the availability of at least one survey item across a selection of eight countries from 1990 to 2017. The other three panels focus on the three most common question themes. We see that the single item data is indeed sparse and that even the aggregated data contains lots of gaps.

```{r load-data}
# read support for dem data. NOTE: We have RespN instead of RespPerc as in online material
df = read.csv("raw-data/supdem raw survey marginals.csv")
```

```{r data-processing}
# remove NAs
df = df[!df$Response==0, ]

# create year by country indicators
year0 = 1987
df = df[df$Year > year0,]
df = unite(df, YearCountry, c(Year, Country), sep = "_", remove = FALSE)

# create item by country indicators
df = unite(df, ItemCnt, c(Item, Country), sep = "_", remove = FALSE)

# create year by project indicators
df = unite(df, YrProj, c(Year, Project), sep = "_", remove = FALSE)

# create year by project by country indicators
df = unite(df, YrProjCnt, c(YrProj, Country), sep = "_", remove = FALSE)


# factorise
df$Country = as.factor(as.character(df$Country))
df$Item = as.factor(as.character(df$Item))
df$ItemCnt = as.factor(as.character(df$ItemCnt))
df$Project = as.factor(as.character(df$Project))
df$YrProj = as.factor(as.character(df$YrProj))
df$YrProjCnt = as.factor(as.character(df$YrProjCnt))
df$Year = df$Year-year0

# count data
dim(df)[1]                   # 3765 national responses
length(unique(df$Country))   # 150 countries
length(unique(df$Project))   # 14 projects
length(unique(df$Item))      # 52 items
length(unique(df$ItemCnt))   # 1453 item-country
length(unique(df$YrProjCnt)) # 1390 national surveys
length(unique(df$Year))      # 27 unique years (out of 30)
sum(df$Sample) / dim(df)[1] * length(unique(df$YrProjCnt))   # 1,804,450 respondents              
```
```{r analyze-years-coverage}
# drop countries with less than 2 years of data
cnt.obs.years = rowSums(table(df$Country, df$Year) > 0)
sort(cnt.obs.years)
mean(cnt.obs.years)
```
```{r drop-countries-with-only-1-year-of-survey-measures}
df2 = df[df$Country %in% levels(df$Country)[cnt.obs.years > 1], ]
length(unique(df2$Country))   # 137 countries with 2+ years of data
```



```{r get-topk-data-countries-number-of-respondants}
sums_respondants_by_country = df2 %>%
    group_by(Country) %>%
    summarize(Sum_Sample = sum(Sample))

topk_sum_respondants = tbl_df(sums_respondants_by_country) %>%
  arrange(sums_respondants_by_country) %>%
  top_n(1)

cond <- unique(topk_sum_respondants$Country)
unique(topk_sum_respondants$Country)

# show data of China as example
filter(df2, Country %in% cond) 
```



```{r sparse-data, label="sparse-data", include=TRUE, echo=FALSE, fig.cap="Sparseness of Aggregate Support for Democacy By Country, Year, and Survey Item. The figure is an update from Claassen (2019).", warning=FALSE, fig.pos="H"}

## Plot indicative grid showing data availability (figure 1)

# plot 1: all items

#pdf("figure1.pdf", width=6.5, height=3.5)
layout(matrix(1:4, nrow=2, ncol=2, byrow=TRUE), widths=c(1.4, 1), heights=c(1, 1.15))

plot.cnt.2 = c('Egypt', 'Venezuela', 'South Africa', 'Botswana', 'China', 'Ukraine', 
               'United States of America', 'Sweden')

plot.dat.2 = df2[df2$Country %in% plot.cnt.2,]
cnt.yr.tab = table(df2$Year, df2$Country)
cnt.yr.tab = ifelse(cnt.yr.tab > 0, 1, NA)
plot.dat.2 = rbind(cnt.yr.tab[,plot.cnt.2[1]], cnt.yr.tab[,plot.cnt.2[2]], cnt.yr.tab[,plot.cnt.2[3]],
                   cnt.yr.tab[,plot.cnt.2[4]], cnt.yr.tab[,plot.cnt.2[5]], cnt.yr.tab[,plot.cnt.2[6]],
                   cnt.yr.tab[,plot.cnt.2[7]], cnt.yr.tab[,plot.cnt.2[8]])
plot.dat.2 = plot.dat.2 * 1:8

par(mar=c(0.3, 7.5, 2, 0.5), tcl=-0.2, cex=0.8)
plot(plot.dat.2[1,], type="n", ylim=c(0.5, 8.5), xaxt="n", yaxt="n", xlab="", ylab="", 
     main='', xlim=c(1, 27), bty="n")
abline(h=1:8, col=grey(0.5), lwd=0.75, lty=3)
abline(v=1:28, col=grey(0.5), lwd=0.75, lty=3)
for(i in 1:8){
  points(plot.dat.2[i,], pch=16, cex=1.1, col='black')
}
axis(2, at=1:8, labels=plot.cnt.2, cex.axis=0.78, mgp=c(1, .5, 0), las=1)
mtext(text='All items', side=3, line=0.4, cex=0.85)
box(lwd=0.75)

# plot 2: only strongman item
cnt.yr.tab = table(df2[grep('Strong', df2$Item), "Year"], 
                   df2[grep('Strong', df2$Item), "Country"])
cnt.yr.tab = ifelse(cnt.yr.tab > 0, 1, NA)
plot.dat.2 = rbind(cnt.yr.tab[,plot.cnt.2[1]], cnt.yr.tab[,plot.cnt.2[2]], cnt.yr.tab[,plot.cnt.2[3]],
                   cnt.yr.tab[,plot.cnt.2[4]], cnt.yr.tab[,plot.cnt.2[5]], cnt.yr.tab[,plot.cnt.2[6]],
                   cnt.yr.tab[,plot.cnt.2[7]], cnt.yr.tab[,plot.cnt.2[8]])
plot.dat.2 = plot.dat.2 * 1:8

par(mar=c(0.3, 0.5, 2, 0.5), tcl=-0.2, cex=0.8)
plot(plot.dat.2[1,], type="p", ylim=c(0.5, 8.5), xaxt="n", yaxt="n", xlab="", ylab="", 
     main='', xlim=c(1, 27), bty="n")
abline(h=1:8, col=grey(0.5), lwd=0.75, lty=3)
abline(v=1:28, col=grey(0.5), lwd=0.75, lty=3)
for(i in 1:8){
  points(plot.dat.2[i,], pch=16, cex=1.1)
}
mtext(text='Strong leader items', side=3, line=0.4, cex=0.85)
box(lwd=0.75)

# plot 3: only three-statements item
cnt.yr.tab = table(df2[grep('ThreeState', df2$Item), "Year"], 
                   df2[grep('ThreeState', df2$Item), "Country"])
cnt.yr.tab = ifelse(cnt.yr.tab > 0, 1, NA)
plot.dat.2 = rbind(cnt.yr.tab[,plot.cnt.2[1]], cnt.yr.tab[,plot.cnt.2[2]], cnt.yr.tab[,plot.cnt.2[3]],
                   cnt.yr.tab[,plot.cnt.2[4]], cnt.yr.tab[,plot.cnt.2[5]], cnt.yr.tab[,plot.cnt.2[6]],
                   cnt.yr.tab[,plot.cnt.2[7]], cnt.yr.tab[,plot.cnt.2[8]])
plot.dat.2 = plot.dat.2 * 1:8

par(mar=c(2, 7.5, 1.8, 0.5), tcl=-0.2, cex=0.8)
plot(plot.dat.2[1,], type="p", ylim=c(0.5, 8.5), xaxt="n", yaxt="n", xlab="", ylab="", 
     main='', xlim=c(1, 27), bty="n")
abline(h=1:8, col=grey(0.5), lwd=0.75, lty=3)
abline(v=1:28, col=grey(0.5), lwd=0.75, lty=3)
for(i in 1:8){
  points(plot.dat.2[i,], pch=16, cex=1.1)
}
axis(1, at=c(4, 9, 14, 19, 24), labels=c(1995, 2000, 2005, 2010, 2015), cex.axis=0.75, mgp=c(1, 0.1, 0))
axis(2, at=1:8, labels=plot.cnt.2, cex.axis=0.78, mgp=c(1, .5, 0), las=1)
mtext(text='Three statements items', side=3, line=0.4, cex=0.85)
mtext(text='Year', side=1, line=1.1, cex=0.7)
box(lwd=0.75)

# plot 4: only army rule item
cnt.yr.tab = table(df2[grep('Army', df2$Item), "Year"], 
                   df2[grep('Army', df2$Item), "Country"])
cnt.yr.tab = ifelse(cnt.yr.tab > 0, 1, NA)
plot.dat.2 = rbind(cnt.yr.tab[,plot.cnt.2[1]], cnt.yr.tab[,plot.cnt.2[2]], cnt.yr.tab[,plot.cnt.2[3]],
                   cnt.yr.tab[,plot.cnt.2[4]], cnt.yr.tab[,plot.cnt.2[5]], cnt.yr.tab[,plot.cnt.2[6]],
                   cnt.yr.tab[,plot.cnt.2[7]], cnt.yr.tab[,plot.cnt.2[8]])
plot.dat.2 = plot.dat.2 * 1:8

par(mar=c(2, 0.5, 1.8, 0.5), tcl=-0.2, cex=0.8)
plot(plot.dat.2[1,], type="p", ylim=c(0.5, 8.5), xaxt="n", yaxt="n", xlab="", ylab="", 
     main='', xlim=c(1, 27), bty="n")
abline(h=1:8, col=grey(0.5), lwd=0.75, lty=3)
abline(v=1:28, col=grey(0.5), lwd=0.75, lty=3)
for(i in 1:8){
  points(plot.dat.2[i,], pch=16, cex=1.1)
}
axis(1, at=c(4, 9, 14, 19, 24), labels=c(1995, 2000, 2005, 2010, 2015), cex.axis=0.75, mgp=c(1, 0.1, 0))
mtext(text='Military rule items', side=3, line=0.4, cex=0.85)
mtext(text='Year', side=1, line=1.1, cex=0.7)
box(lwd=0.75)

#dev.off()



```

\noindent
So far, past research has responded to this issue by discarding most of the available data and using only small cross-sectional data sets consisting of data from only one survey project for one year [@Inglehart2003; @Qi2011; @Welzel2007]. Of course, this leads to a disregard  not only of all other countries but also of additional items. @Claassen2019estimating describes how to use most available data for simulating a dataset that is dense across the yearly time dimension. This dataset can then be used for subsequent statistical analyses [@Claassen2020support; @Claassen2020mood]. The next section describes @Claassen2019estimating's approach.

