---
title: "On Using Metropolis-Hastings to Analyze Democratic Support"
author: "Tobias Stenzel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
header-includes:
   - \usepackage{tikz, float, caption, amsthm, algorithm, algpseudocode}
   - \floatplacement{figure}{H}
   #- \newtheorem{definition}{Definition}
   - \interfootnotelinepenalty=10000 # no multi-page footnotes
output:
  bookdown::pdf_document2:
    template: template.tex
    latex_engine: xelatex
    toc: yes
    toc_depth: 2
    keep_tex: yes
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
bibliography: bibliography.bib
csl-hanging-indent: yes
fontsize: 12pt 
linestretch: 2.0 # adjust for line spacing 
geometry: left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm
classoption:
- a4paper
- oneside
lang: en-EN
numbersections: yes
csquotes: yes
type: Final Paper for Course
course: Advanced Quantitative Methods in Political Science
subtitle: ''
address: ''
email: tobias.stenzel@students.uni-mannheim.de
phone: ''
examiner: Prof. Thomas Gschwend, Ph.D.
chair: ''
mp: 0.55
ID: ''
# STRONGLY BIASED wordcount: '*Wordcount excluding References: `r unima::count_words(knitr::current_input())`*' 
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
editor_options:
  markdown:
    wrap: sentence
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      include=FALSE,
                      fig.path = "figs/",
                      out.width="\\textwidth"
                      )
p_needed <- c(# all packages you need to install here
  "knitr",
  "bookdown", # referencing figures and tables by label
  "remotes",
  "ggplot2",
  "tidyverse", # manipulate df for plotting
  "stargazer",
  "dplyr", # to compute leads values
  "readr", # for csv merge cmd
  "tidyr", # for dropping nans
  "GGally", # correlation matrix style plots
  "gridExtra", # grid for multiple plots
  "cowplot",
  "MASS", # glm
  "janitor", # fd plot over range
  #"countreg", # rootogram
  "ggeasy", # remove axes
  "latex2exp", # latex legend labels
  "ggridges",
  #christopher claassen's additional requirements
  "arm",
  "loo",
  "rstan",
  "RColorBrewer",
  "ggpubr", # for grid in fig
  "directlabels"# label for hline
  )
# installs only the required packages 
lapply(p_needed[!(p_needed %in% rownames(installed.packages()))], install.packages)
lapply(p_needed, library, character.only = TRUE)
# separately install for the correct wordcount package 
if (!"unima" %in% rownames(installed.packages())){
  remotes::install_github("vktrsmnv/unima-template", upgrade = "never", dependencies = TRUE)
  }
# this allows you to add notes to figures with a simple chunk option
# you only need to add "notes="text" as a chunk option; 
# the notes will only appear in PDF output
hook_chunk = knit_hooks$get('chunk')
knit_hooks$set(chunk = function(x, options) {
  txt = hook_chunk(x, options)
  # add chunk option 'description' which adds \Description{...} to figures
  if (!is.null(options$notes)) {
    latex_include <- paste0("\\\\vspace\\{0.5cm\\} \\\\\\footnotesize\\{\\\\textit\\{Notes: \\}", options$notes, "\\} \\1")
    gsub('(\\\\end\\{figure\\})', latex_include, txt) 
  } else {
    return(txt)  # pass to default hook
  }
})
if (knitr::is_latex_output()) knitr::knit_hooks$set(plot = knitr::hook_plot_tex)
# This is an option for stargazer tables
# It automatically adapts the output to html or latex,
# depending on whether we want a html or pdf file
stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")
# This ensures that if the file is knitted to HTML,
# significance notes are depicted correctly 
if (stargazer_opt == "html"){
  fargs <- formals(stargazer)
  fargs$notes.append = FALSE
  fargs$notes = c("<em>&#42;p&lt;0.1;&#42;&#42;p&lt;0.05;&#42;&#42;&#42;p&lt;0.01</em>")
  formals(stargazer) <- fargs
}
# only relevant for ggplot2 plotting
# setting a global ggplot theme for the entire document to avoid 
# setting this individually for each plot 
theme_set(theme_classic() + # start with classic theme 
  theme(
    plot.background = element_blank(),# remove all background 
    plot.title.position = "plot", # move the plot title start slightly 
    legend.position = "bottom" # by default, put legend on the bottom
  ))
```

# Introduction

```{r child = '01-intro.Rmd'}
```



# Literature Review and Data

```{r child = '02-literature-review.Rmd'}
```

# Model and Estimation

```{r child = '03-model-and-estimation.Rmd'}
```

# Analysis and Results 

```{r child = '04-analysis-and-results.Rmd'}
```

# Discussion

```{r child = '05-discussion.Rmd'}
```

# Conclusion

# Appendix: Mathematical Background

```{r child = '07-appendix.Rmd'}
```




\newpage

# References {.unnumbered}

\singlespacing

::: {#refs}
:::


\clearpage

# Statutory Declaration {.unnumbered}

Hiermit versichere ich, dass diese Arbeit von mir persönlich verfasst ist und dass ich keinerlei fremde Hilfe in Anspruch genommen habe.
Ebenso versichere ich, dass diese Arbeit oder Teile daraus weder von mir selbst noch von anderen als Leistungsnachweise andernorts eingereicht wurden.
Wörtliche oder sinngemäße Übernahmen aus anderen Schriften und Veröffentlichungen in gedruckter oder elektronischer Form sind gekennzeichnet.
Sämtliche Sekundärliteratur und sonstige Quellen sind nachgewiesen und in der Bibliographie aufgeführt.
Das Gleiche gilt für graphische Darstellungen und Bilder sowie für alle Internet-Quellen.
Ich bin ferner damit einverstanden, dass meine Arbeit zum Zwecke eines Plagiatsabgleichs in elektronischer Form anonymisiert versendet und gespeichert werden kann.
Mir ist bekannt, dass von der Korrektur der Arbeit abgesehen werden kann, wenn die Erklärung nicht erteilt wird.

```{=tex}
\SignatureAndDate{}
\renewcommand*{\thepage}{ }
```
\noindent I hereby declare that the paper presented is my own work and that I have not called upon the help of a third party.
In addition, I affirm that neither I nor anybody else has submitted this paper or parts of it to obtain credits elsewhere before.
I have clearly marked and acknowledged all quotations or references that have been taken from the works of other.
All secondary literature and other sources are marked and listed in the bibliography.
The same applies to all charts, diagrams and illustrations as well as to all Internet sources.
Moreover, I consent to my paper being electronically stores and sent anonymously in order to be checked for plagiarism.
I am aware that the paper cannot be evaluated and may be graded "failed" ("nicht ausreichend") if the declaration is not made.

\SignatureAndDateEng{}

<!-- Line below depicts the content that should not be counted in the wordcount -->
<!---TC:ignore--->